Carbonate chemistry at Ny-Ålesund

### 2023-03-03 awipev-CO2.Rmd
Air-sea flux in mol C too; refined pangaea file

pangaea <- as_tibble(d_all) %>% 
  dplyr::select(datetime, depth, sal_insitu_ctd, sal_fb, temp_insitu_11m, temp_insitu_ctd, temp_sf, temp_fb, pco2_fb_insi_temp, pHint_tot_sf, at, ct, ph_s_sf_temp_insi) %>%
  dplyr::rename(
    `pressure_profiler [dbar]` = depth,
    # hydrostatic pressure
    `salinity_PSS78_profiler [unit]` = sal_insitu_ctd,
    # salinity in situ
    `salinity_PSS78_ferrybox [unit]` = sal_fb,
    # salinity (FerryBox)
    `temperature_ITS90_11m [°C]` = temp_insitu_11m,
    # temperature in situ (static at 11 m)
    `temperature_ITS90_profiler [°C]` = temp_insitu_ctd,
    # temperature in situ (profiler)
    `temperature_ITS90_ferrybox [°C]` = temp_fb,
    # temperature (FerryBox)
    `temperature_ITS90_seafet_profiler [°C]` = temp_sf,
    # temperature SeaFET (profiler)
    `ta_discrete [umol kg-1]` = at,
    # total alkalinity in situ, discrete measurements
    #`at_calc [umol kg-1]` = at_calc,    # total alkalinity calculated from salinity
    `dic_discrete [umol kg-1]` = ct,
    # dissolved inorganic carbon in situ, discrete measurements
    `ph_insitu_temperature_discrete [total scale]` = ph_s_sf_temp_insi,
    # spectrophotometric pH in situ (total scale) at in situ temperature (discrete)
    #`ph_s_dur_t_fb [total scale]` = ph_s_dur_temp_fb,
    `pco2_insitu_temperature_ferrybox [uatm]` = pco2_fb_insi_temp,
    # partial pressure of CO2 at in situ temperature (FerryBox)
    #`pco2_calc [uatm]` = pco2_calc_fb_insi_temp,
    `ph_insitu_temperature_profiler [total scale]` = pHint_tot_sf,
    # pH in situ at in situ temperature (profiler)
    #`pH_dur [total scale]` = pH_dur,
    #`pH_calc [total scale]` = ph_calc_fb
  ) %>% 
  dplyr::mutate(
    across(starts_with("pressure"), ~ round(.x, digits = 2)),
    across(starts_with("salinity"), ~ round(.x, digits = 2)),
    across(starts_with("temperature"), ~ round(.x, digits = 2)),
    across(starts_with("ph"), ~ round(.x, digits = 3)),
    across(
      starts_with("ta") |
        starts_with("dic") |
        starts_with("pco2"),
      ~ round(.x, digits = 1)
    ),
  )
# replace NaNs by NAs
pangaea <- pangaea %>% 
  mutate_all(~ifelse(is.nan(.), NA, .))

#saveRDS(file= paste0(path, "fb_data/pangaea.rds"), pangaea)
#write_csv(pangaea, file = paste0(path, "fb_data/pangaea.csv"))
saveRDS(file= "data/AWIPEV-CO2_v1.rds", pangaea)
write_csv(pangaea, file = "data/AWIPEV-CO2_v1.csv")
# Metadata Pangaea
# pressure = pressure in situ (profiler)
# s_insitu = salinity in situ (profiler)
# s_fb = salinity (static at 11 m) 
# t_11m = temperature in situ (static at 11 m)
# t_ctd = temperature in situ (profiler)
# t_fb = temperature (FerryBox)
# t_sf = temperature SeaFET (profiler)
# at = total alkalinity, discrete measurements (FerryBox)
# at_calc = total alkalinity, calculated from salinity
# ct = dissolved inorganic carbon, discrete measurements (FerryBox)
# ph_s_sf_t_insi = spectrophotometric pH in situ at in situ temperature (profiler)
# ph_s_dur_t_fb = spectrophotometric pH in FerryBox at FerryBox temperature (FerryBox)
# pco2 = pCO2 in Ferrybox expressed at in situ temperature (FerryBox)
# pco2_calc = pCO2 in the FerryBox calculated from discrete measurements of AT and CT and expressed at in situ temperature
# pH_sf = pH seaFET at in situ temperature (profiler)
# ph_dur = pH Durafet at FerryBox temperature (FerryBox)
# ph_calc = pH at FerryBox temperature calculated from CT and AT in the FerryBox
```

@@

# For paper
size_point <- 1.5 # size of data points
alpha <- 1 #transparency
size_labs <- 14 # font size labels
# temp_insitu_11m
ts_air_sea <- air_sea_month %>%
  ggplot(aes(x = montht, y = f_air_sea / 1000), na.rm = TRUE) +
  geom_col(colour="blue", fill = "blue",  alpha = alpha, stat="identity") +
  lims(y = c(-3, 0)) +
  geom_hline(yintercept = 0) +
  labs(title=NULL, 
       x="Month", 
       y=latex2exp::TeX("Air-sea CO$_2$ flux (mol $m^{-2}$ month$^{-1}$)", bold = TRUE)) +
  Mytheme(size_labs = size_labs) +
  theme(axis.title.x = element_text(face = "bold", size = size_labs))
ts_air_sea
ggsave(filename = "figures/ts_air_sea.png", plot = ts_air_sea, width = 20, units = "cm")
ggsave(filename = "figures/essd/ts_air_sea.pdf", plot = ts_air_sea, width = 20, units = "cm")

#rm(p, p_f_air_sea, p_f_air_sea_d, p_f_air_sea_m)
```

@@

The annual air-sea CO2 flux is
`r round(sum(air_sea_month$f_air_sea/1000), digits = 0)` and
`r round(sum(air_sea_corr_month$f_air_sea_corr/1000), digits = 0)` mol/(m2
year) without and with correction, respectively.

>>>>

pangaea <- as_tibble(d_all) %>% 
  dplyr::select(datetime, depth, sal_insitu_ctd, sal_fb, temp_insitu_11m, temp_insitu_ctd, temp_sf, temp_fb, pco2_fb_insi_temp, pHint_tot_sf, at, ct, ph_s_sf_temp_insi) %>%
  dplyr::rename(
    `pressure_profiler [dbar]` = depth,
    # hydrostatic pressure
    `salinity_PSS78_profiler [unit]` = sal_insitu_ctd,
    # salinity in situ
    `salinity_PSS78_ferrybox [unit]` = sal_fb,
    # salinity (FerryBox)
    `temperature_ITS90_11m [°C]` = temp_insitu_11m,
    # temperature in situ (static at 11 m)
    `temperature_ITS90_profiler [°C]` = temp_insitu_ctd,
    # temperature in situ (profiler)
    `temperature_ITS90_ferrybox [°C]` = temp_fb,
    # temperature (FerryBox)
    `temperature_ITS90_seafet_profiler [°C]` = temp_sf,
    # temperature SeaFET (profiler)
    `ta_discrete [umol kg-1]` = at,
    # total alkalinity in situ, discrete measurements
    #`at_calc [umol kg-1]` = at_calc,    # total alkalinity calculated from salinity
    `dic_discrete [umol kg-1]` = ct,
    # dissolved inorganic carbon in situ, discrete measurements
    `ph_insitu_temperature_discrete [total scale]` = ph_s_sf_temp_insi,
    # spectrophotometric pH in situ (total scale) at in situ temperature (discrete)
    #`ph_s_dur_t_fb [total scale]` = ph_s_dur_temp_fb,
    `pco2_insitu_temperature_ferrybox [uatm]` = pco2_fb_insi_temp,
    # partial pressure of CO2 at in situ temperature (FerryBox)
    #`pco2_calc [uatm]` = pco2_calc_fb_insi_temp,
    `ph_insitu_temperature_profiler [total scale]` = pHint_tot_sf,
    # pH in situ at in situ temperature (profiler)
    #`pH_dur [total scale]` = pH_dur,
    #`pH_calc [total scale]` = ph_calc_fb
  ) %>% 
  dplyr::mutate(
    across(starts_with("pressure"), ~ round(.x, digits = 2)),
    across(starts_with("salinity"), ~ round(.x, digits = 2)),
    across(starts_with("temperature"), ~ round(.x, digits = 2)),
    across(starts_with("ph"), ~ round(.x, digits = 3)),
    across(
      starts_with("ta") |
        starts_with("dic") |
        starts_with("pco2"),
      ~ round(.x, digits = 1)
    ),
  )
# replace NaNs by NAs, then remove NAs
pangaea <- pangaea %>% 
  mutate_all(~ifelse(is.nan(.), NA, .)) %>% 
  mutate(datetime = as.POSIXct(datetime, origin = "1970-01-01")) |>
  separate(datetime, into = c("Date", "Time"), sep = " ") |>
  mutate(`date/time [UTC+0]` = paste(Date, Time, sep = "T")) |>
  dplyr::select(`date/time [UTC+0]`, everything(), -Date, -Time) |>
  mutate_at(1:13, ~as.character(.)) |>
  mutate_at(1:13, ~replace_na(., ""))

saveRDS(file= "data/AWIPEV-CO2_v1.rds", pangaea)
write_csv(pangaea, file = "data/AWIPEV-CO2_v1.csv")
# Metadata Pangaea (see manuscript)
# datetime (UTC)
# pressure = pressure in situ (profiler)
# s_insitu = salinity in situ (profiler)
# s_fb = salinity (static at 11 m) 
# t_11m = temperature in situ (static at 11 m)
# t_ctd = temperature in situ (profiler)
# t_fb = temperature (FerryBox)
# t_sf = temperature SeaFET (profiler)
# at = total alkalinity, discrete measurements (FerryBox)
# at_calc = total alkalinity, calculated from salinity
# ct = dissolved inorganic carbon, discrete measurements (FerryBox)
# ph_s_sf_t_insi = spectrophotometric pH in situ at in situ temperature (profiler)
# ph_s_dur_t_fb = spectrophotometric pH in FerryBox at FerryBox temperature (FerryBox)
# pco2 = pCO2 in Ferrybox expressed at in situ temperature (FerryBox)
# pco2_calc = pCO2 in the FerryBox calculated from discrete measurements of AT and CT and expressed at in situ temperature
# pH_sf = pH seaFET at in situ temperature (profiler)
# ph_dur = pH Durafet at FerryBox temperature (FerryBox)
# ph_calc = pH at FerryBox temperature calculated from CT and AT in the FerryBox
```

@@

# For paper
size_point <- 1.5 # size of data points
alpha <- 1 #transparency
size_labs <- 14 # font size labels
# temp_insitu_11m
ts_air_sea <- air_sea_month %>%
  ggplot(aes(x = montht, y = f_air_sea / 1000), na.rm = TRUE) +
  geom_col(colour="blue", fill = "blue",  alpha = alpha, stat="identity") +
  lims(y = c(-3, 0)) +
  geom_hline(yintercept = 0) +
  labs(title=NULL, 
       x="Month", 
       y=latex2exp::TeX("Air-sea CO$_2$ flux (mol $m^{-2}$ month$^{-1}$)", bold = TRUE)) +
  Mytheme(size_labs = size_labs) +
  theme(axis.title.x = element_text(face = "bold", size = size_labs))
ts_air_sea
ggsave(filename = "figures/ts_air_sea.png", plot = ts_air_sea, width = 20, units = "cm")
ggsave(filename = "figures/essd/ts_air_sea.pdf", plot = ts_air_sea, width = 20, units = "cm")
#rm(p, p_f_air_sea, p_f_air_sea_d, p_f_air_sea_m)
```

@@

The annual air-sea CO2 flux is
`r round(sum(air_sea_month$f_air_sea/1000), digits = 1)` and
`r round(sum(air_sea_corr_month$f_air_sea_corr/1000), digits = 1)` mol CO2/(m2
year) without and with correction, respectively. Or `r round(sum(air_sea_month$f_air_sea*(12/44)/1000), digits = 1)` and
`r round(sum(air_sea_corr_month$f_air_sea_corr*(12/44)/1000), digits = 1)` mol C/(m2
year)
